Bayesian Nonparametric System Reliability\\ using Sets of Priors

### 2016-01-20 ijar-npb-ms.tex
New results in Section 6 (and proofs in appendix) on first-order stochastic dominance to avoid numerical optimisation for y and n.

@@

To obtain the lower and upper bound for the system reliability function $\Rsys(t)$,
we now need to minimise and maximise Equation~\eqref{eq:rsyswithsurvsign} over $\PtZi{1}, \ldots, \PtZi{K}$ for each $t$,
where the posterior predictive probabilities for $C^k_t$ are given by the Beta-Binomial pmf \eqref{eq:postpredCny}.
It turns out that we need to optimise over $\ntz{1}, \ldots, \ntz{K}$ only, %$\nktz$ ($k=1,\ldots,K$) only,
as lower and upper bounds for $\Rsys(t)$ are obtained for $\yktzl$ and $\yktzu$, respectively.

To see this,
remember that $\yktz$ gives the prior expected functioning probability for a component of type $k$ at time $t$.
From Equation~\eqref{eq:nyupdate} (right) we see that $\yktn$ is increasing in $\yktz$,
%the posterior parameter set bounds are the transformed prior parameter set bounds,
%i.e., $\yktzu$ corresponds to the upper bound of $\PktN$, for all $\nktz \in \PktZ$,
such that for each fixed $\nktz$, $\yktzu$ will maximise the posterior expected functioning probability $\yktn$.
%
Now, a higher posterior expected functioning probability means
higher probability for many components out of $m_k$ surviving,
i.e., probabilities for the larger $l_k$'s are higher, and in turn for the lower $l_k$'s are lower.
Indeed, the Beta-Binomial distribution is stochastically ordered for varying $\yktn$, i.e.,
its cmf \eqref{eq:postpredCnycmf} is decreasing in $\yktn$ for all $l_k$,
such that the cmfs for different values of $\yktn$ do not cross each other.
This means that for each fixed $\nktz$, the Beta-Binomial cmf \eqref{eq:postpredCnycmf} is minimized for $\yktzu$ in all $l_k$,
such that the probability weights given to high values of $l_k$ are maximised. %for $\yktzu$.
For coherent systems, $\Phi$ is non-decreasing in each of its arguments $l_1,\ldots,l_K$,
thus $\Rsys(t)$ is maximised by putting most probability weight on the highest $l_k$ values for each $k$,
which is in turn obtained for $\yktzu$, $k=1,\ldots,K$.
We therefore have
%
%for a certain $k$, a high probability on high $l_k$'s is never worse than a high probability on lower $l_k$'s,
%keeping all other component type expressions fixed.
%***Monotonicity in $\yz_{k,t}$, so we get, for each $t \in {\cal T}$,
\begin{align}
\lefteqn{\lRsys(t \mid \vec{t}^1, \ldots, \vec{t}^K)} \nonumber \\
 &= \min_{\PtZi{1}, \ldots, \PtZi{K}} \Rsys(t \mid \PtZi{1}, \ldots, \PtZi{K}, \vec{t}^1, \ldots, \vec{t}^K) \nonumber \\
 &= \min_{\nz_{1,t},\ldots,\nz_{K,t}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K)
                                                 \prod_{k=1}^K P(C^k_t = l_k \mid \yktzl, \nktz, s^k_t) \nonumber \\
 &= \min_{\nz_{1,t},\ldots,\nz_{K,t}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \nonumber \\ & \hspace*{12ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nn_{k,t}\ynl_{k,t}, m_k - l_k + \nn_{k,t}(1-\ynl_{k,t}))}
                                         {B(\nn_{k,t}\ynl_{k,t}, \nn_{k,t}(1-\ynl_{k,t}))} \,. \label{eq:LwrSysPost}
\end{align}
In the following, we will give some guidelines on how to choose the parameter sets $\PkZi{1}, \ldots, \PkZi{\tmax}$
which define the set of prior discrete reliability functions for components of type $k$.
We think this is much easier in terms of $\nz$ and $\yz$ than it would be in terms of $\alpha$ and $\beta$.
%***How to choose the sets of priors for the survival function of type $k$ components,
%i.e. how to choose $\PkZi{1}, \ldots \PkZi{\tmax}$?***

@@

\section*{Appendix}
\appendix
\renewcommand*{\thesection}{\Alph{section}}
\section{Software details}
\label{ap:software}

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

@@

The elements reviewed and extended above culminate hereinafter in the primary contribution of the current work,
providing a framework in which the nonparametric Bayesian system reliability approach developed in \cite{2015:bayessurvsign}
is extended to incorporate the sets of priors approach of \citet{2009:WalterAugustin},
enabling diagnoses of prior-data conflict which is consequential at the system
level, via sets of system reliability functions.

To obtain the lower and upper bound for the system reliability function $\Rsys(t)$,
we now need to minimise and maximise Equation~\eqref{eq:rsyswithsurvsign} over $\PtZi{1}, \ldots, \PtZi{K}$ for each $t$,
where the posterior predictive probabilities for $C^k_t$ are given by the Beta-Binomial pmf \eqref{eq:postpredCny}.
%It turns out that we need to optimise over $\ntz{1}, \ldots, \ntz{K}$ only, %$\nktz$ ($k=1,\ldots,K$) only,
%as lower and upper bounds for $\Rsys(t)$ are obtained for $\yktzl$ and $\yktzu$, respectively.
%
%To see this,
%remember that $\yktz$ gives the prior expected functioning probability for a component of type $k$ at time $t$.
%From Equation~\eqref{eq:nyupdate} (right) we see that $\yktn$ is increasing in $\yktz$,
%%the posterior parameter set bounds are the transformed prior parameter set bounds,
%%i.e., $\yktzu$ corresponds to the upper bound of $\PktN$, for all $\nktz \in \PktZ$,
%such that for each fixed $\nktz$, $\yktzu$ will maximise the posterior expected functioning probability $\yktn$.
%%
%Now, a higher posterior expected functioning probability means
%higher probability for many components out of $m_k$ surviving,
%i.e., probabilities for the larger $l_k$'s are higher, and in turn for the lower $l_k$'s are lower.
%Indeed, the Beta-Binomial distribution is stochastically ordered for varying $\yktn$, i.e.,
%its cmf \eqref{eq:postpredCnycmf} is decreasing in $\yktn$ for all $l_k$,
%such that the cmfs for different values of $\yktn$ do not cross each other.
%This means that for each fixed $\nktz$, the Beta-Binomial cmf \eqref{eq:postpredCnycmf} is minimized for $\yktzu$ in all $l_k$,
%such that the probability weights given to high values of $l_k$ are maximised. %for $\yktzu$.
%For coherent systems, $\Phi$ is non-decreasing in each of its arguments $l_1,\ldots,l_K$,
%thus $\Rsys(t)$ is maximised by putting most probability weight on the highest $l_k$ values for each $k$,
%which is in turn obtained for $\yktzu$, $k=1,\ldots,K$.
We therefore have
%
%for a certain $k$, a high probability on high $l_k$'s is never worse than a high probability on lower $l_k$'s,
%keeping all other component type expressions fixed.
%***Monotonicity in $\yz_{k,t}$, so we get, for each $t \in {\cal T}$,
\begin{align}
\lefteqn{\lRsys(t \mid \vec{t}^1, \ldots, \vec{t}^K)} \nonumber \\
 &= \min_{\PtZi{1}, \ldots, \PtZi{K}} \Rsys(t \mid \PtZi{1}, \ldots, \PtZi{K}, \vec{t}^1, \ldots, \vec{t}^K) \nonumber \\
 &= \min_{\PtZi{1}, \ldots, \PtZi{K}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K)
                                                 \prod_{k=1}^K P(C^k_t = l_k \mid \yktz, \nktz, s^k_t) \nonumber \\
 &= \min_{\PtZi{1}, \ldots, \PtZi{K}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \nonumber \\ & \hspace*{12ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nn_{k,t}\yn_{k,t}, m_k - l_k + \nn_{k,t}(1-\yn_{k,t}))}
                                         {B(\nn_{k,t}\yn_{k,t}, \nn_{k,t}(1-\yn_{k,t}))} \nonumber \\
 &= \min_{\PtZi{1}, \ldots, \PtZi{K}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \nonumber \\ & \hspace*{8ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nktz\yktz + s^k_t, m_k - l_k + \nktz(1-\yktz) + n_k - s^k_t)}{B(\nktz\yktz + s^k_t, \nktz(1-\yktz) + n_k - s^k_t)}
    \,, \label{eq:LwrSysPostA}
\end{align}
and similarly maximising for $\uRsys(\cdot)$.

Note that $\Phi(\cdot)$ is non-decreasing in each of its arguments $l_1,\ldots,l_K$,
thus if there is first-order stochastic ordering on $P(C^k_t = l_k \mid \yktz, \nktz, s^k_t)$
for each $k$, then this ordering can be used to determine the elements of $\PtZi{k}$
which minimise and maximise the overall system reliability function without
resorting to computationally expensive exhaustive searches or numerical optimisation.

We therefore start by proving the following result, where indices are suppressed for readability.
We use $\ge_{\mathrm{st}}$ to denote first-order stochastic dominance.

\begin{theorem}
  \label{thm:y}
  Let $\beta_y$ denote the Beta-Binomial distribution with probability mass function parameterised as:
  \[ p(l \mid y, n, m, s, N) \propto \frac{B(l + ny + s, m - l + n(1-y) + N - s)}{B(ny + s, n(1-y) + N - s)}, \]
  with $n, m, s,$ and $N$ fixed and unknown.

  Then $\beta_{\yu} \ge_{\mathrm{st}} \beta_{\yl} \ \forall \ \yu > \yl$ with $\yu, \yl \in (0,1)$.
\end{theorem}
The proof is provided in Appendix \ref{ap:proofs}, p.\pageref{prf:y}.

Consequently, for each component, the posterior predictive Beta Binomial distributions 
with larger prior functioning probability stochastically dominate those
with smaller prior functioning probability, providing rigorous proof 
which accords with intuition. Applying this result to the sets of system reliability 
functions, together with the monotonicity in the survival signature, means that
$\lRsys(\cdot)$ is attained when $\yktz = \yktzl$ and $\uRsys(\cdot)$
is attained when $\yktz = \yktzu$ for all possible $\nktz$ values.

The analogous result for $\nktz$ is more subtle, because stochastic 
dominance is not guaranteed at a single value.  The following Theorem 
provides the conditions under which an upper or lower limit has first-order stochastic dominance.

\begin{theorem}
  \label{thm:n}
  Let $\beta_n$ denote the Beta-Binomial distribution with probability mass function parameterised as:
  \[ p(l \mid y, n, m, s, N) \propto \frac{B(l + ny + s, m - l + n(1-y) + N - s)}{B(ny + s, n(1-y) + N - s)}, \]
  with $y, m, s,$ and $N$ fixed and unknown.  
  Then,
  \[ y(N+m-1)-s-m > 0 \implies \beta_{\nu} \ge_{\mathrm{st}} \beta_{\nl} \]
  and 
  \[ y(N+m-1)-s-m < 0 \implies \beta_{\nu} \le_{\mathrm{st}} \beta_{\nl} \]
\end{theorem}
The proof is provided in Appendix \ref{ap:proofs}, p.\pageref{prf:n}.

These first-order stochastic domination results now gives the exact 
solution for $\lRsys(\cdot)$ and $\uRsys(\cdot)$.  Thus,
%For $\lRsys(\cdot)$, the minimum is achieved at
%$\nktzu$ if $\yktzl(n_k+m_k-1)-s^k_t-m_k < 0$ and is achieved at 
%$\nktzl$ if $\yktzl(n_k+m_k-1)-s^k_t-m_k > 0$.
%Conversely, for $\uRsys(\cdot)$, the maximum is achieved at 
%$\nktzu$ if $\yktzu(n_k+m_k-1)-s^k_t-m_k > 0$ and is achieved at 
%$\nktzl$ if $\yktzu(n_k+m_k-1)-s^k_t-m_k < 0$.
\begin{align}
\lefteqn{\lRsys(t \mid \vec{t}^1, \ldots, \vec{t}^K)} \nonumber \\
 &= \min_{\PtZi{1}, \ldots, \PtZi{K}} 
    \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \nonumber \\ & \hspace*{8ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nktz\yktz + s^k_t, m_k - l_k + \nktz(1-\yktz) + n_k - s^k_t)}{B(\nktz\yktz + s^k_t, \nktz(1-\yktz) + n_k - s^k_t)}
    \nonumber\\
 &= \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} \Phi(l_1, \ldots, l_K) \times \nonumber \\ & \hspace*{8ex}
    \prod_{k=1}^K {m_k \choose l_k} \frac{B(l_k + \nktzo\yktzl + s^k_t, m_k - l_k + \nktzo(1-\yktzl) + n_k - s^k_t)}{B(\nktzo\yktzl + s^k_t, \nktzo(1-\yktzl) + n_k - s^k_t)} \label{eq:LwrSysPost}\\
 &\qquad \qquad \mbox{where } \nktzo = \left\{ \begin{aligned}
   \nktzu & \quad\mbox{if}\quad \yktzl(n_k+m_k-1)-s^k_t-m_k < 0 \\
   \nktzl & \quad\mbox{if}\quad \yktzl(n_k+m_k-1)-s^k_t-m_k > 0 \\
 \end{aligned} \right. \nonumber
\end{align}

\td{*** Louis: I've edited to here so far, will continue editing rest of this section if we are happy with the above! ***}

In the following, we will give some guidelines on how to choose the parameter sets $\PkZi{1}, \ldots, \PkZi{\tmax}$
which define the set of prior discrete reliability functions for components of type $k$.
We think this is much easier in terms of $\nz$ and $\yz$ than it would be in terms of $\alpha$ and $\beta$.
%***How to choose the sets of priors for the survival function of type $k$ components,
%i.e. how to choose $\PkZi{1}, \ldots \PkZi{\tmax}$?***

@@

\section*{Appendix}
\appendix
\renewcommand*{\thesection}{\Alph{section}}

\section{Proofs}
\label{ap:proofs}

\begin{proof}[\textbf{Proof of Theorem \ref{thm:y}, p\pageref{thm:y}}]
  \label{prf:y}
  Consider the likelihood ratio for the two Beta Binomial distributions $\beta_{\yu}$ and $\beta_{\yl}$,
  \begin{align*}
    \lefteqn{\mathcal{L}(l) := \frac{p(l \mid \yu, n, m, s, N)}{p(l \mid \yl, n, m, s, N)}} \\
    &= \frac{B(l + n \yu + s, m - l + n (1 - \yu) + N - s) B(n \yl + s, n (1 - \yl) + N - s)}
            {B(n \yu + s, n (1 - \yu) + N - s) B(l + n \yl + s, m - l + n (1 - \yl) + N - s)} \\
    &= \frac{\Gamma(l + n \yu + s) \Gamma(m - l + n (1 - \yu) + N - s) \Gamma(n \yl + s) \Gamma(n (1 - \yl) + N - s)}
            {\Gamma(l + n \yl + s) \Gamma(m - l + n (1 - \yl) + N - s) \Gamma(n \yu + s) \Gamma(n (1 - \yu) + N - s)} \\
    &= \left\{ \begin{aligned}
         \frac{\prod_{x=0}^{m-1} (x + n (1 - \yu) + N - s)}
              {\prod_{x=0}^{m-1} (x + n (1 - \yl) + N - s)} &\quad\mbox{ for } l=0 \\
         \frac{\prod_{x=0}^{l-1} (x + n \yu + s) \prod_{x=0}^{m-l-1} (x + n (1 - \yu) + N - s)}
              {\prod_{x=0}^{l-1} (x + n \yl + s) \prod_{x=0}^{m-l-1} (x + n (1 - \yl) + N - s)} &\quad\mbox{ for } 0<l<m \\
         \frac{\prod_{x=0}^{m-1} (x + n \yu + s)}
              {\prod_{x=0}^{m-1} (x + n \yl + s)} &\quad\mbox{ for } l=m
       \end{aligned} \right.
  \end{align*}
  since $\Gamma(x+1)=x \Gamma(x)$.

  Thus,
  \begin{align*}
    \frac{\mathcal{L}(l+1)}{\mathcal{L}(l)} &=
      \frac{(l + n \yu) (m - l - 1 + n (1 - \yl) + N - s)}
           {(l + n \yl) (m - l - 1 + n (1 - \yu) + N - s)} \\
    &> 1 \quad\mbox{when}\quad 0 \le \yl < \yu \le 1
  \end{align*}

  Hence, $\mathcal{L}(\cdot)$ is monotone increasing for $0 < \yl < \yu < 1$, so that $\beta_{\yu}$ is larger than or equal to $\beta_{\yl}$ in monotone likelihood ratio order ($\beta_{\yu} \ge_\mathrm{lr} \beta_{\yl}$).  But, $\beta_{\yu} \ge_\mathrm{lr} \beta_{\yl} \implies \beta_{\yu} \ge_\mathrm{st} \beta_{\yl}$ (\cite[Theorem 1.C.1, p.43]{shaked2007}) giving the required result.
\end{proof}

\begin{proof}[\textbf{Proof of Theorem \ref{thm:n}, p\pageref{thm:n}}]
  \label{prf:n}
  Consider the likelihood ratio for the two Beta Binomial distributions $\beta_{\nu}$ and $\beta_{\nl}$,
  \begin{align*}
    \lefteqn{\mathcal{L}(l) := \frac{p(l \mid y, \nu, m, s, N)}{p(l \mid y, \nl, m, s, N)}} \\
    &= \frac{B(l + \nu y + s, m - l + \nu (1 - y) + N - s) B(\nl y + s, \nl (1 - y) + N - s)}
            {B(\nu y + s, \nu (1 - y) + N - s) B(l + \nl y + s, m - l + \nl (1 - y) + N - s)} \\
    &= \frac{\Gamma(l + \nu y + s) \Gamma(m - l + \nu (1 - y) + N - s)}
            {\Gamma(l + \nl y + s) \Gamma(m - l + \nl (1 - y) + N - s)} \\
    &\quad \times \frac{\Gamma(\nl y + s) \Gamma(\nl (1 - y) + N - s) \Gamma(\nu+N) \Gamma(m+\nl+N)}
                       {\Gamma(\nu y + s) \Gamma(\nu (1 - y) + N - s)\Gamma(\nl+N) \Gamma(m+\nu+N)} \\
    &= \left\{ \begin{aligned}
         \frac{\prod_{x=0}^{m-1} (x + \nu (1 - y) + N - s) \prod_{x=0}^{m-1} (x + \nl + N)}
              {\prod_{x=0}^{m-1} (x + \nl (1 - y) + N - s) \prod_{x=0}^{m-1} (x + \nu + N)} &\quad\mbox{ for } l=0 \\
         \frac{\prod_{x=0}^{l-1} (x + \nu y + s) \prod_{x=0}^{m-l-1} (x + \nu (1 - y) + N - s)}
              {\prod_{x=0}^{l-1} (x + \nl y + s) \prod_{x=0}^{m-l-1} (x + \nl (1 - y) + N - s)} \\
         \quad \times \frac{\prod_{x=0}^{m-1} (x + \nl + N)}{\prod_{x=0}^{m-1} (x + \nu + N)} &\quad\mbox{ for } 0<l<m \\
         \frac{\prod_{x=0}^{m-1} (x + \nu y + s) \prod_{x=0}^{m-1} (x + \nl + N)}
              {\prod_{x=0}^{m-1} (x + \nl y + s) \prod_{x=0}^{m-1} (x + \nu + N)} &\quad\mbox{ for } l=m
       \end{aligned} \right.
  \end{align*}
  since $\Gamma(x+1)=x \Gamma(x)$.

  Thus,
  \[ \frac{\mathcal{L}(l+1)}{\mathcal{L}(l)} =
      \frac{(l + \nu y) (m - l - 1 + \nl (1 - y) + N - s)}
           {(l + \nl y) (m - l - 1 + \nu (1 - y) + N - s)}
  \]

  However, unlike the case for the $y$ parameter in Theorem \ref{thm:y}, neither $\beta_{\nl}$ 
  nor $\beta_{\nu}$ can be guaranteed to dominate for all possible values 
  for the other parameters, so that necessary conditions for
  monotonicity (either increasing or decreasing) must be established.
  We require,
  \[ \frac{(l + \nu y) (m - l - 1 + \nl (1 - y) + N - s)}
          {(l + \nl y) (m - l - 1 + \nu (1 - y) + N - s)}
     >1
  \]
  After extensive routine algebra, this can be conveniently expressed as
  \begin{align*}
    (\nu - \nl)[y (N + m) - s - y] - l (\nu - \nl)>0.
  \end{align*}
  This limit is hardest to satisfy for $l=m$ since $\nu-\nl>0$.
  Thus, for monotonicity to hold for all $l$, we require
  \begin{align*}
    (\nu - \nl)[y (N + m) - s - y] - m (\nu - \nl) &>0 \\
    \implies (\nu - \nl)[y (N + m - 1) - s - y] &>0
  \end{align*}
  Since $\nu-\nl>0$ by definition, we have a monotonically increasing 
  likelihood ratio only when \[ y (N + m - 1) - s - y > 0. \]  Likewise 
  the likelihood ratio is only monotonically decreasing when
  \[ y (N + m - 1) - s - y < 0. \]
  Thus,
  \[ y(N+m-1)-s-m > 0 \implies \beta_{\nu} \ge_{\mathrm{lr}} \beta_{\nl} \implies \beta_{\nu} \ge_{\mathrm{st}} \beta_{\nl} \]
  and 
  \[ y(N+m-1)-s-m < 0 \implies \beta_{\nu} \le_{\mathrm{lr}} \beta_{\nl} \implies \beta_{\nu} \le_{\mathrm{st}} \beta_{\nl} \]
  by \cite[Theorem 1.C.1, p.43]{shaked2007}.
\end{proof}

\section{Software details}
\label{ap:software}
